{ \fontfamily{ptm}\selectfont Catalyst Meta Acceleration Framework: The history and the gist of it }

### 2024-11-26 Notebook LaTEX/Catalyst Accelerations/main.tex
>w<, changes

\section{Introduction}
    \textcolor{red}{THIS REPORT IS CURRENTLY: FINISHED, BUT NOT YET POLISHED}

@@

\section{Lin's Estimating Sequence}\label{sec:lin-2015}
    The section introduces the Nesterov's estimating sequence in Lin 2015 \cite{lin_universal_2015}. 
    We warn the readers about the followings: 
    \begin{enumerate}
        \item The proofs in HongZhou Lin's original paper of Universal Catalyst is depressingly long and complicated. It is a result of using the constructive approach of Nesterov's estimating sequence. 
        \item Controlling the errors of inexact proximal point evaluations is context specific. Lin hinted at ways to track the errors such as using duality and non-negativity assumption of the objective. He illustrated the use of the meta acceleration on their own method called: ``Accelerated MISO-Prox", in general there is not a universal solution. 
        \item We will provide proofs to clarify some of their proofs and compare with existing proofs and drawing references in the literatures in the appendix. 
    \end{enumerate}
    Let's assume $F$ is a $\mu \ge 0$ strongly convex function. 
    Throughout this section we make the following notations
    \begin{align*}
        \mathcal M^{\kappa^{-1}}(x; y) &:= F(x) + \frac{\kappa}{2}\Vert x - y\Vert^2, 
        \\
        \mathcal J_{\kappa^{-1}} y &:= \argmin_x \mathcal M^{\kappa^{-1}} (x, y). 
    \end{align*}
    Their algorithm is almost exactly the same as Nesterov's 2.2.20 \cite{nesterov_lectures_2018} which we stated in the definition below: 
    \begin{definition}[Lin's accelerated proximal point method]
        Define the estimating sequence $(\phi_k)_{k \ge0}$ recursively by: 
        \begin{align*}
            \phi_0(x) &:= F(x_0) + \frac{\gamma_0}{2}\Vert x - v_0\Vert^2, 
            \\
            \phi_k(x) &:= 
            (1 - \alpha_{k - 1})\phi_{k - 1}(x)
            + \alpha_{k - 1}\left(
                F(x_k) + \kappa\langle y_{k - 1} - x_k, x - x_k \rangle
                + \frac{\mu}{2}\Vert x - x_k\Vert^2
            \right). 
        \end{align*}
        Let the initial estimate be $x_0 \in \RR^n$, fix parameters $\kappa$ and $\alpha_0$. 
        Let $(\epsilon_k)_{k \ge 0}$ be an error sequence chosen for the evaluation for inexact proximal point method. 
        Initialize $x_0 = y_0$, then the algorithm generates $(x_k, y_k)_{k\ge 0}$ for all $k \ge 1$ such that: 
        \begin{align*}
            & \text{find } x_k \approx \mathcal J_{\kappa^{-1}} y_{k - 1} \text{ such that } \mathcal M^{\kappa^{-1}}(x_k, y_{k - 1}) - \mathcal M^{\kappa^{-1}}(\mathcal J_{\kappa^{-1}}y_{k - 1}, y_{k - 1}) \le \epsilon_k
            \\
            & \text{find } \alpha_k \in (0, 1) \text{ such that } \alpha_k^2 = (1 - \alpha_k)\alpha_{k - 1}^2 + (\mu/(\mu + \kappa)) 
            \\
            & 
            y_{k} = x_k + \frac{\alpha_{k - 1}(1 - \alpha_{k - 1})}{\alpha_{k - 1}^2 + \alpha_k}(x_k - x_{k - 1}). 
        \end{align*}
    \end{definition}
    \begin{remark}
        The algorithm is similarity to
        \hyperref[def:nes-est-seq-pg]{Definition \ref*{def:nes-est-seq-pg}}. 
        In contrast, it has an inexact proximal point step controlled by $\epsilon_k$, and the Lipschitz constant $L$ is absent instead it has $\kappa + \mu$. 
        Evaluating $x_k \approx \mathcal J_{\kappa^{-1}}y_{k - 1}$ is also possible because the function $\mathcal M^{\kappa^{-1}}(\cdot, y_{k - 1})$ is strongly convex, hence its optimality gap can be bounded via trackable quantity $\partial \mathcal M^{\kappa^{-1}}(x_k, y_{k - 1})$.

>>>>

\section{Introduction}
    \textcolor{red}{THIS REPORT IS CURRENTLY: FIRST POLISHING DONE. }

@@

\section{Lin's estimating sequence}\label{sec:lin-2015}
    The section introduces the Nesterov's estimating sequence in Lin 2015 \cite{lin_universal_2015}. 
    We warn the readers about the followings: 
    \begin{enumerate}
        \item The proofs in HongZhou Lin's original paper of Universal Catalyst is depressingly long and complicated. It is a result of using the constructive approach of Nesterov's estimating sequence. 
        \item Controlling the errors of inexact proximal point evaluations is context specific. Lin hinted at ways to track the errors such as using duality and non-negativity assumption of the objective. He illustrated the use of the meta acceleration on their own method called: ``Accelerated MISO-Prox", in general there is not a universal solution. 
        \item We will provide proofs to clarify some of their proofs and compare with existing proofs and drawing references in the literatures in the appendix. 
    \end{enumerate}
    Let's assume $F$ is a $\mu \ge 0$ strongly convex function. 
    Throughout this section we make the following notations
    \begin{align*}
        \mathcal M^{\kappa^{-1}}(x; y) &:= F(x) + \frac{\kappa}{2}\Vert x - y\Vert^2, 
        \\
        \mathcal J_{\kappa^{-1}} y &:= \argmin_x \mathcal M^{\kappa^{-1}} (x, y). 
    \end{align*}
    Their algorithm is almost exactly the same as Nesterov's 2.2.20 \cite{nesterov_lectures_2018} which we stated in the definition below: 
    \begin{definition}[Lin's accelerated proximal point method]
        Define the estimating sequence $(\phi_k)_{k \ge0}$ recursively by: 
        \begin{align*}
            \phi_0(x) &:= F(x_0) + \frac{\gamma_0}{2}\Vert x - v_0\Vert^2, 
            \\
            \phi_k(x) &:= 
            (1 - \alpha_{k - 1})\phi_{k - 1}(x)
            + \alpha_{k - 1}\left(
                F(x_k) + \kappa\langle y_{k - 1} - x_k, x - x_k \rangle
                + \frac{\mu}{2}\Vert x - x_k\Vert^2
            \right). 
        \end{align*}
        Let the initial estimate be $x_0 \in \RR^n$, fix parameters $\kappa$ and $\alpha_0$. 
        Let $(\epsilon_k)_{k \ge 0}$ be an error sequence chosen for the evaluation for inexact proximal point method. 
        Initialize $x_0 = y_0$, then the algorithm generates $(x_k, y_k)_{k\ge 0}$ for all $k \ge 1$ such that: 
        \begin{align*}
            & \text{find } x_k \approx \mathcal J_{\kappa^{-1}} y_{k - 1} \text{ such that } \mathcal M^{\kappa^{-1}}(x_k, y_{k - 1}) - \mathcal M^{\kappa^{-1}}(\mathcal J_{\kappa^{-1}}y_{k - 1}, y_{k - 1}) \le \epsilon_k
            \\
            & \text{find } \alpha_k \in (0, 1) \text{ such that } \alpha_k^2 = (1 - \alpha_k)\alpha_{k - 1}^2 + (\mu/(\mu + \kappa)) 
            \\
            & 
            y_{k} = x_k + \frac{\alpha_{k - 1}(1 - \alpha_{k - 1})}{\alpha_{k - 1}^2 + \alpha_k}(x_k - x_{k - 1}). 
        \end{align*}
    \end{definition}
    \begin{remark}
        The algorithm is similarity to
        \hyperref[def:nes-est-seq-pg]{Definition \ref*{def:nes-est-seq-pg}}. 
        In contrast, it has an inexact proximal point step controlled by $\epsilon_k$, and the Lipschitz constant $L$ is absent instead it has $\kappa + \mu$. 
        Evaluating $x_k \approx \mathcal J_{\kappa^{-1}}y_{k - 1}$ is also possible because the function $\mathcal M^{\kappa^{-1}}(\cdot, y_{k - 1})$ is strongly convex, hence its optimality gap can be bounded via trackable quantity $\partial \mathcal M^{\kappa^{-1}}(x_k, y_{k - 1})$.
